---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# rrcov3way: Robust Methods for Multiway Data Analysis

[![CRAN version](https://www.r-pkg.org/badges/version/rrcov3way)](https://cran.r-project.org/package=rrcov3way)
[![downloads](https://cranlogs.r-pkg.org/badges/rrcov3way)](https://cran.r-project.org/package=rrcov3way)

The package provides methods for multiway data analysis by means of _Parafac_ and _Tucker 3_ models. Robust versions (Engelen and Hubert (2011) <doi:10.1016/j.aca.2011.04.043>) and versions for compositional data are also provided (Gallo (2015) <doi:10.1080/03610926.2013.798664>, Di Palma et al. (2018) <doi:10.1080/02664763.2017.1381669>).

## Installation

The `rrcov3way` package is on CRAN (The Comprehensive R Archive Network) and
the latest release can be easily installed using the command

    install.packages("rrcov3way")

## Building from source

To install the latest stable development version from GitHub, you can
pull this repository and install it using

    ## install.packages("remotes")
    remotes::install_github("valentint/rrcov3way")

Of course, if you have already installed `remotes`, you can skip the
first line (I have commented it out).

## Example

This is a simple example which shows you basic functions of the package using the OECD `elind` data set. The data consist of specialization indices of electronics industries of 23 European countries for the years 1973-1979. The specialization index is defined as the proportion of the monetary value of an electronic industry compared to the total export value of manufactured goods of a country compared to the similar proportion for the world as a whole. 

```{r example}

## Load the package 'rrcov3way' and the data set
##  to be used in the examples:
##
## OECD Electronics Industries Data - Kroonenberg PM (2008)
##  23 countries x 6 industries x 7 years
library(rrcov3way)
data(elind)
dim(elind)

##  The frontal slices (mode C) are matrices with dimension 23x6, 
##  representing the data for all countries and all industries 
##  in one year. The labels are:
rownames(elind[,,1])
colnames(elind[,,1])

##  The lateral slices (mode B) represent the data
##  for all countries and years for one industry
##  - matrices with dimension 23 x 7 and the labels are:
rownames(elind[,1,])
colnames(elind[,1,])

##  First of all we center and scale the data, using the default procedures 
##  for centering and scaling.

elind <- do3Scale(elind, center=TRUE, scale=TRUE)

##  Next we perform classical PARAFAC analysis 
##  with the default number of components (ncomp=2).

## default PARAFAC, non-robust, no ilr transformation,
##  extract 3 components
res <- Parafac(elind, ncomp=3)
res


```
